---
title: "Part2_Regression_Model"
author: "Sharika Loganathan"
date: "2022-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages

```{r, load_packages}
library(tidyverse)
library(bayesplot)
library(rstanarm)
library(coefplot)
library(splines)


```

## Read data


```{r, read_glimpse_data}
df <- readr::read_csv("fall2022_finalproject.csv", col_names = TRUE)
```

```{r}
head(df)
```
```{r}
df %>% glimpse()
```

You are required as part of the project to explore the data. Your exploration will demonstrate that `output`, the continuous response, is between 0 and 1. Because of this, it is **highly recommended** that you transform the continuous response before training regression models. You should use the logit transformation to convert the lower and upper bounded `output` variable to an unbounded variable. The regression models should be trained to predict the logit-transformed response. The code chunk below shows how to calculate the unbounded response, `y`, as the logit transformation of the `output` variable.  

```{r}
df_reg <- df %>% 
  
  mutate(
    x5 = 1 - (x1 + x2 + x3 + x4),
         w = x2 / (x3 + x4),
         z = (x1 + x2) / (x5 + x4),
         t = v1 * v2,
         y = boot::logit(output)
    ) %>% 
  select(x1,x2,x3,x4,v1,v2,v3,v4,v4,v5,x5,w,z,t,m,y) %>% 
  glimpse()
```


## Models using the “base feature”

### All linear additive features

```{r}
mod_01 <- lm(y ~ x1+x2+x3+x4+v1+v2+v3+v4+v5+m , data = df_reg)
broom::glance(mod_01)
coefplot::coefplot(mod_01)
```


###  Interaction of the categorical input with all continuous inputs


```{r}
mod_02 <- lm(y ~ m *(x1 + x2 + x3 + x4+v1+v2+v3+v4+v5) , data = df_reg)
broom::glance(mod_02)
coefplot::coefplot(mod_02)
```



###   All pair-wise interactions of the continuous inputs

```{r}
mod_03 <- lm( y ~ (x1 + x2 + x3 + x4+v1+v2+v3+v4+v5)^2, data = df_reg)
broom::glance(mod_03)
coefplot::coefplot(mod_03)
```





##  Models using the “expanded feature”/ Derived  feature

### Linear additive features


```{r}
mod_04 <- lm( y ~ (x5+w+t+z+m), data = df_reg)
broom::glance(mod_04)
coefplot::coefplot(mod_04)
```

###   Interaction of the categorical input with continuous features

```{r}
mod_05 <- lm( y ~ m*(x5+w+t+z), data = df_reg)
broom::glance(mod_05)
coefplot::coefplot(mod_05)
```

###   Pair-wise interactions between the continuous features



```{r}
mod_06 <- lm( y ~ m*(x5+w+t+z)^2, data = df_reg)
broom::glance(mod_06)
coefplot::coefplot(mod_06)
```



##    Models linear basis function models


###  All linear additive features

```{r}
mod_07 <- lm(y~ (.+splines::ns(x1,df=3) +splines::ns(x5,df=3) +splines::ns(z,df=3) -x1 -x5 -z), data=df_reg)
broom::glance(mod_07)
coefplot::coefplot(mod_07)
```


###   Interaction of the categorical input with all continuous inputs

Combining chemistry inputs with manufacturing processing units and process

```{r}

mod_08 <- lm(y~ (m)*(splines::ns(x5, 3)  + splines::ns(w, 3) + splines::ns(z, 3)), data=df_reg)
broom::glance(mod_08)
coefplot::coefplot(mod_08)
```
##

```{r}


mod_09<- lm(y~ m+ (splines::ns(x1, 3) * splines::ns(x3, 3) * splines::ns(x5, 2)),data=df_reg)
broom::glance(mod_09)
coefplot::coefplot(mod_09)
```


```{r}
mod_10<- lm(y~ m*(splines::ns(x5, 3) * splines::ns(w, 3) * splines::ns(z, 3)),data=df_reg)
broom::glance(mod_10)
coefplot::coefplot(mod_10)
```


##  Train 10 different models

##   Performance Metrics


```{r}
extract_metrics <- function(mod, mod_name)
{
  broom::glance(mod) %>% 
    mutate(model_name = mod_name)
}

model_results <- purrr::map2_dfr(list(mod_01, mod_02, mod_03, mod_04,
                                      mod_05, mod_06, mod_07, mod_08,mod_09,mod_10
                                      ),
                                 sprintf("mod-%02d", 1:10),
                                 extract_metrics)

```



```{r}
model_results %>% 
  select(r.squared, model_name) %>% 
  arrange(desc(r.squared))
```


Model 7 is the best model according to r-squared.





```{r}
model_results %>% 
  ggplot(mapping = aes(x = model_name, y = r.squared)) +
  geom_linerange(mapping = aes(ymin = 0,
                               ymax = r.squared)) +
  geom_point(size = 4.5) +
  labs(x = '') +
  theme_bw()
```

Which of the 9 models is the best

mod 7 is the best .

What performance metric did you use to make your selection?

```{r}
model_results %>% 
  select(model_name, r.squared, AIC, BIC) %>% 
  pivot_longer(!c("model_name")) %>% 
  mutate(model_id = stringr::str_extract(model_name, "\\d+")) %>% 
  ggplot(mapping = aes(x = model_id, y = value)) +
  geom_point(size = 3.5) +
  facet_wrap(~name, scales = "free_y") +
  labs(x = '') +
  theme_bw()
```

Model 7 is the best model.

Visualize the coefficient summaries for your top 3 models.

Mod_07 ,Mod_09, Mod_10 are the top 3 models.

```{r}
coefplot::multiplot(mod_07,mod_09,mod_10)
```


How do the coefficient summaries compare between the top 3 models?

coefficient of mod 7 is more 


```{r}
mod_07 %>% readr::write_rds('mod_07.rds')
mod_08 %>% readr::write_rds('mod_08.rds')
mod_10 %>% readr::write_rds('mod_10.rds')
```



##   Regression– iiB) Bayesian Linear models




```{r}
re_load_mod_07 <- readr::read_rds('mod_07.rds')
```






stan_lm for base and derived inputs 

```{r}
set.seed(43212)
bayes_mod01 <- stan_lm( y~ m+ (splines::ns(x1, 2) * splines::ns(x3, 2) * splines::ns(x5, 2)), data = df_reg,
                 prior = rstanarm::R2(location = 0.5),
                 seed = 432123)

bayes_mod02 <- stan_lm( y ~ (.+splines::ns(x1,df=2) +splines::ns(x5,df=2) +splines::ns(z,df=2) -x1 -x5 -z), data = df_reg,
                 prior = rstanarm::R2(location = 0.5),
                 seed = 432123)


bayes_mod03 <- stan_lm( y~ (.+splines::ns(x1,df=3) +splines::ns(x5,df=3) +splines::ns(z,df=3) -x1 -x5 -z), data = df_reg,
                 prior = rstanarm::R2(location = 0.5),
                 seed = 432123)


```

```{r}
bayes_mod01 %>% summary()
bayes_mod02 %>% summary()
bayes_mod03 %>% summary()
```

```{r}
posterior_interval(bayes_mod01)
posterior_interval(bayes_mod02)
posterior_interval(bayes_mod03)
```

```{r}
rstanarm::bayes_R2(bayes_mod01) %>% quantile(c(0.05, 0.5, 0.95))
```




###   Posterior visualizations


```{r}
plot(bayes_mod01) + theme_bw()

```




```{r}
plot(bayes_mod02) + theme_bw()
```



Alternatively, you may use rstanarm’s stan_lm() or stan_glm()
function to fit full Bayesian linear models with syntax like R’s lm().


```{r}
plot(bayes_mod03) + theme_bw()
```





After fitting the 2 models, you must identify the best model. Which performance metric did you use to make your selection?


```{r}
all_models_rsq <-purrr::map2_dfr(list(bayes_mod01, bayes_mod02 ,bayes_mod03),
                as.character(1:3),
                function(mod, mod_name){tibble::tibble(rsquared = bayes_R2(mod)) %>% 
                    mutate(model_name = mod_name)}) 
all_models_rsq%>% 
  ggplot(mapping = aes(x = rsquared)) +
  geom_freqpoly(bins = 55,
                 mapping = aes(color = model_name),
                 size = 1.1) +
  coord_cartesian(xlim = c(0, 1)) +
  ggthemes::scale_color_colorblind("Model") +
  theme_bw()
```

```{r}
purrr::map2_dfr(list(bayes_mod01, bayes_mod02,bayes_mod03),
                as.character(1:3),
                function(mod, mod_name){as.data.frame(mod) %>% tibble::as_tibble() %>% 
                    select(sigma) %>% 
                    mutate(model_name = mod_name)}) %>% 
  ggplot(mapping = aes(x = sigma)) +
  geom_freqpoly(bins = 55,
                 mapping = aes(color = model_name),
                 size = 1.1) +
  ggthemes::scale_color_colorblind("Model") +
  theme_bw()
```

```{r}
all_models_rsq %>%
  ggplot(mapping = aes(x = model_name, y = rsquared, fill = model_name)) +
  geom_boxplot(width = 0.3) + 
  scale_fill_brewer(palette = "Set2") +
  theme_bw() +
  theme(aspect.ratio = 1)
```


Performance of both the models are almost same, so I choose model 1 as it’s performance is slightly better.



Visualize the regression coefficient posterior summary statistics for your best model.


##   Visualizing Posterior coefficient summary for the best model


```{r}
plot(bayes_mod01, pars = names(bayes_mod01$coefficients)) +
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed", size = 1.0) +
  theme_bw()
```






##  Read Mod_07

```{r}
re_load_mod_07 <- readr::read_rds('mod_07.rds')
```


##   Regression – iiC) Linear models Predictions


```{r}
as.data.frame(bayes_mod03) %>% tibble::as_tibble() %>% 
  ggplot(mapping = aes(x = sigma)) +
  geom_histogram(bins = 55) +
  theme_bw()
```


```{r}
as.data.frame(bayes_mod03) %>% tibble::as_tibble() %>% 
  ggplot(mapping = aes(x = sigma)) +
  geom_histogram(bins = 55) +
  geom_vline(xintercept = stats::sigma(re_load_mod_07), 
             color = "red", linetype = "dashed", size = 1.1)
```


```{r}
as.data.frame(bayes_mod01) %>% tibble::as_tibble() %>% 
  select(sigma) %>% 
  pull() %>% 
  quantile(c(0.05, 0.5, 0.95))
```


```{r}
bayes_mod01 %>% readr::write_rds('bayes_mod01.rds')
bayes_mod02 %>% readr::write_rds('bayes_mod02.rds')
bayes_mod03 %>% readr::write_rds('bayes_mod03.rds')
```




For your best model: Study the posterior uncertainty in the noise (residual error), 𝜎. How does the lm() maximum likelihood estimate (MLE) on 𝜎 relate to the posterior uncertainty on 𝜎?
• Do you feel the posterior is precise or are we quite uncertain about 𝜎?


###  Testing

```{r}
bayes_mod01 <- readr::read_rds('bayes_mod01.rds')
bayes_mod02 <- readr::read_rds('bayes_mod02.rds')
bayes_mod03 <- readr::read_rds('bayes_mod03.rds')
```




###  Holdout set Prediction

```{r}
holdout <- readr::read_csv('fall2022_holdout_inputs.csv', col_names = TRUE)
```

```{r}
df_holdout <- holdout %>% 
  mutate(x5 = 1 - (x1 + x2 + x3 + x4),
         w = x2 / (x3 + x4),
         z = (x1 + x2) / (x5 + x4),
         t = v1 * v2) %>% 
         select(x1,x2,x3,x4,v1,v2,v3,v4,v4,v5,x5,w,z,t,m) %>%
  glimpse()
```

```{r}
df_test_reg <- df_reg %>% 
  select(x1,x2,x3,x4,v1,v2,v3,v4,v5,x5,w,z,t,m) %>% 
  glimpse()
```


```{r}
sprintf("columns in df_all: %d vs columns in holdout: %d", ncol(df_test_reg), ncol(df_holdout))
```


```{r}
df_holdout %>% names()
```

```{r}
df_test_reg %>% names()
```



```{r}
pred_01 <- posterior_predict(bayes_mod01,df_holdout) 

```


```{r}
pred_02 <- posterior_predict(bayes_mod02, df_holdout) 
```

```{r}
pred_03 <- posterior_predict(bayes_mod03, df_holdout) 
```



##  Regression – iiD) Train/tune with resampling

Train, assess, tune, and compare more complex methods via resampling.


Caret  to handle the preprocessing, training, testing, and evaluation.


###   Resampling and performance metrics



```{r}

library(caret)

my_ctrl <- trainControl(method = "repeatedcv", number = 5, repeats = 3)

my_metric <- "RMSE"
```


##  Basic linear models

```{r}
set.seed(2021)

fit_lm_1 <- train(y ~ x1 + x2 + x3 + x4+v1+v2+v3+v4+v5+m,
                  data = df_reg,
                  method = "lm",
                  metric = my_metric,
                  preProcess = c("center", "scale"),
                  trControl = my_ctrl)

fit_lm_1
```


```{r}
set.seed(2021)

fit_lm_2 <- train(y ~ x5+t+z+w,
                  data = df_reg,
                  method = "lm",
                  metric = my_metric,
                  preProcess = c("center", "scale"),
                  trControl = my_ctrl)

fit_lm_2
```

#   Regularized regression with elastic net

##  Regression – iiD) Train/tune with resampling


Train and tune the neural network, random forest, and the gradient boosted tree with the “base feature” set AND AGAIN with the “expanded feature” set.

Regularized regression with Elastic net
• Interact the categorical variable with all pair-wise interactions of the continuous features.
• The more complex of the 2 models selected from iiA)


```{r}
set.seed(2021)

fit_enet_1 <- train(y ~ (x5+z+t+w)^2 +I(x5^2) + I(z^3),
                    data = df_reg,
                    method = "glmnet",
                    metric = my_metric,
                    preProcess = c("center", "scale"),
                    trControl = my_ctrl)

fit_enet_1
```

Elastic net with all pair-wise interactions between all 5 inputs and polynomial features for log_f.


```{r}
set.seed(2021)

fit_enet_3 <- train(y ~ m*( x5 + z + t+w)^2 + I(x5^2) + I(z^3),
                    data = df_reg,
                    method = "glmnet",
                    metric = my_metric,
                    preProcess = c("center", "scale"),
                    trControl = my_ctrl)

fit_enet_3
```

Neural network – once with “base feature” set & once with “expanded feature” set

##Neural network


```{r}
set.seed(2021)

fit_nnet_1 <- train(y ~ m*(x5+t+z+w),
                    data = df_reg,
                    method = "nnet",
                    metric = my_metric,
                    preProcess = c("center", "scale"),
                    trControl = my_ctrl,
                    trace = FALSE,
                    linout = TRUE)
```

```{r}
fit_nnet_1
```


Let’s try to tune the neural network to see if performance can be improved.

```{r}
nnet_grid <- expand.grid(size = c(2, 4, 6, 8, 10, 12),
                         decay = exp(seq(-6, 2, length.out = 13)))

set.seed(2021)

fit_nnet_2 <- train(y ~ m*(x5+t+w+z),
                    data = df_reg,
                    method = "nnet",
                    metric = my_metric,
                    tuneGrid = nnet_grid,
                    preProcess = c("center", "scale"),
                    trControl = my_ctrl,
                    trace = FALSE,
                    linout = TRUE)
```


Visualize the tuning results.

```{r}
plot(fit_nnet_2, xTrans = log)
```


```{r}
fit_nnet_2$bestTune
```


Random forest – once with “base feature” set & once with “expanded feature” set


#Random forest

```{r}
set.seed(2021)

fit_rf <- train(y ~ m*(x5+t+w+z),
                data = df_reg,
                method = "rf",
                metric = my_metric,
                trControl = my_ctrl,
                importance = TRUE)

fit_rf
```



Gradient boosted tree – once with “base feature” set & once with “expanded feature” set

##Gradient boosted tree
####There are multiple implementations of gradient boosted trees. We will use XGBoost.

```{r}
set.seed(2021)

fit_xgb <- train(y ~ m*(x5+w+t+z),
                 data = df_reg,
                 method = "xgbTree",
                 metric = my_metric,
                 trControl = my_ctrl,
                 objective = 'reg:squarederror')

fit_xgb$bestTune
```


```{r}
plot(fit_xgb)
```
Tuning

```{r}
xgb_grid <- expand.grid(nrounds = seq(100, 700, by = 100),
                        max_depth = c(3, 4, 5),
                        eta = c(0.5*fit_xgb$bestTune$eta, fit_xgb$bestTune$eta),
                        gamma = fit_xgb$bestTune$gamma,
                        colsample_bytree = fit_xgb$bestTune$colsample_bytree,
                        min_child_weight = fit_xgb$bestTune$min_child_weight,
                        subsample = fit_xgb$bestTune$subsample)

set.seed(2021)

fit_xgb_tune <- train(y ~ m*(x5+w+t+z),
                      data = df_reg,
                      method = "xgbTree",
                      tuneGrid = xgb_grid,
                      metric = my_metric,
                      trControl = my_ctrl,
                      objective = 'reg:squarederror')
```


```{r}
plot(fit_xgb_tune)
```


##   Compare models
##   Compile the resampling results together.


```{r}
my_results <- resamples(list(LM_1 = fit_lm_1,
                             LM_2 = fit_lm_2,
                             ENET_1 = fit_enet_1,
                             
                             ENET_3 = fit_enet_3,
                        
                             NNET = fit_nnet_1,
                             NNET_2 = fit_nnet_2,
                          
                           
                             RF = fit_rf,
                             XGB = fit_xgb_tune))
```


```{r}
#Compare models based on RMSE.

dotplot(my_results, metric = "RMSE")
```


```{r}
#Compare the models based on R-squared.

dotplot(my_results, metric = "Rsquared")
```

Which inputs seem important?

###   Variable importances

```{r}
plot(varImp(fit_rf))
```


```{r}
plot(varImp(fit_xgb_tune))
```







Decide the resampling scheme, what kind of pre processing options you should consider, and the performance metric you will focus on.




Identifying the best model.


XGboost is the best model.


z,x5,w, t are the most important features.


```{r}
fit_xgb %>% readr::write_rds('reg_best_model.rds')

```




